Risk Perceptions for Wearable Devices

### 2015-03-06 tex-inputs/related_work.tex

\section{Related Work}
%%% better introduction, or take this out. This seems awkward. 
%We discuss related work that has examined users' perceptions surrounding security and privacy risks.

\subsection{Ubiquity and Wearable Devices}
Many authors have emphasized that we are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Many researchers have worked to study how privacy can be preserved in the presence ubiquitous devices. Examples of such efforts include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy}, protocols for anonymous communication {cornelius2008anonysense}, evaluation metrics for privacy ~\cite{scholtz2004toward}, and privacy models ~\cite{hong2004privacy, jiang2002approximate}. However, none of these works attempted to quantify or rank user concern over different privacy risks.

Roesner et al.\ urge the community to address potential concerns for wearable devices before the technologies become widespread \cite{roesner2014security} and explore the unique and difficult problems these devices present in terms for law and policy \cite.{roesner2014augmented}. A small-scale interview of how bystanders feel about wearable devices found that bystanders were predominantly split between having indifferent and negative reactions to the device but expressed clear interest in being able to give permissions for the data being captured \cite{denning2014situ}. We hope our research furthers   work in communicating with users and bystanders.

\subsection{Smartphones}
Many researchers have attempted to study user concerns about security or privacy issues associated with their smartphones. Research shows that perceptions of risk change as computing systems change. Chin et. al\ investigate how user attitudes toward security and privacy for smarphones differ from attitudes for traditional computing systems, and find that these attitudes differ significantly ~\cite{chin2012measuring}. How people use these devices had also changed. Palen et. al\ track smartphone users for six weeks to find rapid modifications in behaviors with respect to context and the social appropriateness of phone use cite\{palen2000going}. People also perceive and behave without a realistic understanding of the risks they are taking. Felt et. al\ examine the Android permission system and find that 17\% of participants paid attention to permissions, and only 3\% comprehended what these permissions meant.

%%% There's way more work in this area that needs to be cited (look at stuff by Janne Lindqvist, Norman Sadeh, Patrick Kelley, and Jason Hong). At the very least, if you don't add new citations, you absolutely need to say 2-3 sentences about the three papers that you did cite. Though I strongly urge you to add more.

\subsection{User Perceptions and Behaviors}
In this paper we focus on measuring people's perceptions of security and privacy risks. One limitation of user perceptions is that people don't always have enough information to make privacy-sensitive decisions; even if they do, they often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Also, actual behavior may deviate from self-reported behaviors~\cite{jensen2005privacy} and privacy preferences~\cite{spiekermann2001privacy}.

While risk communication for the physical world has been examined for several decades (e.g.,~\cite{Fischhoff,Morgan2001}), research into effectively communicating computer-based risks has only recently been researched. For example, both Garg et al.\ and Blythe et al.\ show that due to varying perceptions and abilities that correlate with demographic factors, computer-based risk communication should employ some degree of demographic targeting~\cite{Garg2012,Blythe2011}. While this work is likely applicable to wearable computing risk communication, we believe that a better understanding of users' risk perceptions in this domain is warranted prior to examining risk communication.

>>>>

\section{Related Work}
%%% better introduction, or take this out. This seems awkward. 
%We discuss related work that has examined users' perceptions surrounding security and privacy risks.
In this section, we outline previous research that has examined privacy for ubiquitous and wearable computing environments, smartphone privacy and security, as well as user risk perceptions and behaviors.

\subsection{Ubiquitous and Wearable Devices}
Many authors have emphasized that we are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Many researchers have worked to study how privacy can be preserved in the presence of ubiquitous devices. Examples of such efforts include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy}, protocols for anonymous communication~\cite{cornelius2008anonysense}, evaluation metrics for privacy~\cite{scholtz2004toward}, and privacy models~\cite{hong2004privacy, jiang2002approximate}. However, none of these works attempted to quantify or rank user concern over different privacy risks.

Roesner {\it et al.}\ urge the community to address potential concerns for wearable devices before the technologies become widespread~\cite{roesner2014security} and explore the unique and difficult problems these devices present in terms for law and policy~\cite{roesner2014augmented}. A small-scale interview of how bystanders feel about wearable devices found that bystanders were predominantly split between having indifference and negative reactions to the device, but expressed clear interest in being able to give permission for data to be captured~\cite{denning2014situ}. Our research furthers this goal of communicating risks to users and bystanders.

\subsection{Smartphones}
Many researchers have attempted to study user concerns about security or privacy issues associated with their smartphones. Research shows that perceptions of risk change based on the particular device being used. Chin {\it et al.}\ found that users' attitudes toward security and privacy for smarphones significantly differed from attitudes towards traditional computing systems~\cite{chin2012measuring}. How people used their smartphones also differed from how they used other computing systems. Palen {\it et al.}\ tracked smartphone users for six weeks and observed radical changes with respect to how they used their devices~\cite{palen2000going}. People also perceive and behave without a realistic understanding of the risks they are taking. Felt {\it et al.}\ examined the Android permission system and found that 17\% of participants paid attention to permission requests, and only 3\% comprehended what these permissions meant.

Smartphones allow applications to access new types of data. While this tends to benefit users, they do not think of the privacy implications. Lindqvist {\it et al.}\  studied a popular location tracking application, and found that people share their location information for gaming, signaling availability to friends, without concern for the privacy implications of broadcasting that information~\cite{lindqvist2011m}. Tsai {\it et al.}\ found that when mobile users get feedback about releasing data, such as who has viewed location information, it greatly impacts future behaviors~\cite{Tsai2009}. However, this type of feedback is not generally provided to users. There are still many unresolved concerns, such as the opaqueness that prevents users from fully understanding how applications are using their data or rogue applications inappropriately accessing data~\cite{1_kane_2010, zhou2011taming}.

\subsection{User Perceptions and Behaviors}
In this paper, we focus on measuring user perceptions of security and privacy risks, as they relate to wearable devices. One limitation of user perceptions is that people do not always have enough information to make privacy-sensitive decisions; even if they do, they often trade off long-term privacy for short-term benefits~\cite{acquisti2005privacy}. Also, actual behavior may deviate from self-reported behaviors~\cite{jensen2005privacy} and stated privacy preferences~\cite{spiekermann2001privacy}.

While risk communication for the physical world has been examined for several decades (e.g.,~\cite{Fischhoff,Morgan2001}), research into effectively communicating computer-based risks has only recently been researched. For example, both Garg {\it et al.}\ and Blythe {\it et al.}\ show that due to varying perceptions and abilities that correlate with demographic factors, computer-based risk communication should employ some degree of demographic targeting~\cite{Garg2012,Blythe2011}. While this work is likely applicable to wearable computing risk communication, we believe that a better understanding of users' risk perceptions in this domain is warranted, prior to examining risk communication.
